### 2020-04-09 paper/paper.tex

\documentclass[twocolumn,11pt]{IEEEtran}
\usepackage[verbose,expansion=alltext,stretch=50]{microtype}
\usepackage{graphicx}

\title{DeepSched: A Deep Representation Of Scheduling Policies For Heterogeneous Distributed Systems}

\author{
   \begin{tabular}{c| c| c| c}
       Mohamed Shawky & Remonda Talaat & Mahmoud Adas & Evram Youssef\\
       \texttt{\small{SEC 2, B.N 16}} & \texttt{\small{SEC 1, B.N 20}} & \texttt{\small{SEC 2, B.N 21}} & \texttt{\small{SEC 1, B.N 9}}
   \end{tabular}%

   \texttt{\small{\{mohamedshawky911, remondatalaat21, mido3ds, evramyousef\}@gmail.com}}
}%

\markboth{Cairo Uni. CMP Dep. $|$ OS Research $|$ Final Paper $|$ \today}{Shell}

\begin{document}
\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
\input{core/introduction}

\section{Related Work}
\input{core/related_work}

\section{Methodology}
\input{core/methodology}

\section{Experimental Setup}
\input{core/exp_setup}

\section{Discussion and Results}
\input{core/results}

\section{Conclusion}
\input{core/conclusion}

\medskip

\bibliographystyle{unsrt}
\bibliography{paper}

\end{document}