### 2020-04-10 paper/core/results.tex
added results chart and table

\subsection{Scheduling Performance}

\begin{figure}[hp]
    \centering
    \includegraphics[width=0.45\textwidth]{diagrams/ta_chart}
    \caption{Average turnaround time produced from the three discussed schedulers at different input sizes.}
    \label{fig:ta_chart}
\end{figure}

\subsection{Execution Time}

\begin{table}
  \caption{Time comparison between the three discussed schedulers at different input sizes. Time measured in seconds.}
  \label{tab:exec_time}
  \centering
  \begin{tabular}{l l l l}
    \toprule
    Input Size & HEFT & GA & NN \\
    \midrule
    486	& 31.5 & 67.5 & 0.125 \\
    10000 & 305.5 & 602.5 & 0.125 \\
    84654 & 670.0 & 3967.0 & 0.125 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Results Discussion}